#Install Library
```{r}
source("C:\\Research\\TP Analysis\\Library\\Library.R")
```

#Data
```{r}
Remote_Data <- read_feather("C:\\Research\\TP Analysis\\Data\\Remote Data\\CLEAR-TP.feather") #33497
NHD_Shape <- st_read("C:\\Research\\TP Analysis\\Shape File\\NHDShapeFile\\nhdplusv2_modified_v1.0.shp")
```

#Directory for dam data
```{r}
Dam_NID <- read.csv("C:\\Research\\TP Analysis\\Data\\Dam Data\\NHD_100_NID.csv") # Dams 
```

#Averaging the remote TP data (Annual TP)
```{r}
Remote_Data <- Remote_Data%>%mutate(Date= as.Date(Date))%>%mutate(Year = year(Date), Month = month(Date))%>%
  mutate(Season = case_when(
    Month %in% 3:5 ~ "Spring",
    Month %in% 6:8 ~ "Summer",
    Month %in% 9:11 ~ "Autumn",
    TRUE ~ "Winter" ))

Remote_Data <- Remote_Data%>%mutate(ID_Year = paste(ID, Year, sep = "_"))

#By Season
Remote_Season_ID <- Remote_Data %>%
  group_by(ID, Year) %>%
  summarise(Season_Count = n_distinct(Season)) %>%
  filter(Season_Count > 3) %>%
  mutate(ID_Year = paste(ID, Year, sep = "_")) # Criteria to consider year will all seasons

Remote_Filtered_Data <- Remote_Data %>%
  filter(ID_Year %in% Remote_Season_ID$ID_Year)

#Averaging
Remote_Data_Average <- Remote_Filtered_Data%>%group_by(ID, Year, Season)%>%summarise(TP_Season = mean(Estimated))
Remote_Data_Average_Year <- Remote_Data_Average%>%group_by(ID, Year)%>%summarise(TP = mean(TP_Season))

Remote_Data_Average_Year <- Remote_Data_Average_Year%>%filter(Year <2019)
```

#Find the location of dams
```{r}
Dam_NID <- Dam_NID%>%mutate(YEAR = Year.Completed, DAM_HGT_M = Dam.Height..feet.* 0.3048, GRAND_ID = River.or.Stream )
Dam_NID <- Dam_NID%>%select(c("GRAND_ID", "ID", "YEAR", "DAM_HGT_M", "Primary.Purpose"))

Dam <- rbind(Dam_NID)
Dam <- Dam%>%group_by(ID)%>%summarise(YEAR = max(YEAR), DAM_HGT_M = mean(DAM_HGT_M))
Dam <- Dam%>%filter(DAM_HGT_M >30.5)
```

#Join the Dam with the river network to defind the upstream and the downstream reach
```{r}
Dam_NHD <- Dam %>%
  left_join(NHD_Shape, by = "ID") %>%
  mutate_all(~ ifelse(is.na(.), 0, .))%>%as.data.frame()

Dam_name <- unique(Dam$ID)
```

#Compare the TP between the reaches
```{r}
Dam_change <- data.frame()
result<- data.frame()
Dam_name <- unique(Dam$ID)

for (j in Dam_name){

      Dam_specific <- j
      
      print(Dam_specific)
      
      Dam_NHD_GNIS <- Dam_NHD %>%
              filter(ID == Dam_specific) %>%
              pull(GNIS_ID)
      
          
          # Get the construction year as an integer
      Cons_year <- Dam %>%
              filter(ID == Dam_specific) %>%
              select(YEAR) %>%
              pull(YEAR)%>%max()
          
      Dam_height <- Dam%>%filter(ID == Dam_specific)%>%select(DAM_HGT_M)%>%pull(DAM_HGT_M)
          
      # Filter remote data for years greater than 5 years after the construction year
      Remote_averaged_filtered <- Remote_Data_Average_Year %>%
              filter(Year > Cons_year)
          
      Dam_NHD_GNIS <- Dam_NHD %>%
              filter(ID == Dam_specific) %>%
              pull(GNIS_ID)
      
      Upstream_reach_ID_1 <- NHD_Shape %>%
        filter(toID == Dam_specific & GNIS_ID == Dam_NHD_GNIS) %>%
        pull(ID)
      
      # Check if the first upstream reach ID exists
      if (length(Upstream_reach_ID_1) == 0 || Upstream_reach_ID_1 == 0) {
        Upstream_reach_ID_2 <- 0
        Upstream_reach_ID_3 <- 0
      } else {
        # Get the second upstream reach ID
        Upstream_reach_ID_2 <- NHD_Shape %>%
          filter(toID == Upstream_reach_ID_1 & GNIS_ID == Dam_NHD_GNIS) %>%
          pull(ID)
        
        # Check if the second upstream reach ID exists
        if (length(Upstream_reach_ID_2) == 0 || Upstream_reach_ID_2 == 0) {
          Upstream_reach_ID_3 <- 0
        } else {
          # Get the third upstream reach ID
          Upstream_reach_ID_3 <- NHD_Shape %>%
            filter(toID == Upstream_reach_ID_2 & GNIS_ID == Dam_NHD_GNIS) %>%
            pull(ID)
        }
      }
      
      # Create a list of the upstream reach IDs
      upstream_ids <- list(Upstream_reach_ID_1, Upstream_reach_ID_2, Upstream_reach_ID_3)
      
      if (all(unlist(upstream_ids) == 0)) {
        print("All values in upstream_ids are 0. Breaking the loop!")
        next
      }
      
      # Initialize a list to store the filtered dataframes
      upstream_years <- list()
      
      # Loop over each Upstream reach ID
      for (i in 1:3) {
        # Get the current upstream reach ID
        reach_id <- upstream_ids[[i]]
        
        # Check if the reach ID is 0, then replace with NA
        if (is_empty(reach_id) || is.na(reach_id) || reach_id == "") {
          # Replace with NA if reach ID is 0
          upstream_years[[i]] <- Remote_averaged_filtered[0, ] %>%
            rename(!!paste0("TP_", i, "_Up") := TP)  # Dynamically name the TP column
        } else {
          # Otherwise, filter and rename
          upstream_years[[i]] <- Remote_averaged_filtered %>%
            filter(ID == reach_id) %>%
            rename(!!paste0("TP_", i, "_Up") := TP)  # Dynamically name the TP column
        }
      }
      
      # Assign each upstream year dataframe to the appropriate variable
      Upstream_year_1 <- upstream_years[[1]]
      Upstream_year_2 <- upstream_years[[2]]
      Upstream_year_3 <- upstream_years[[3]]
      
      #_____________________________________________________________________________   
      
      Dam_year <- Remote_averaged_filtered %>% filter(ID == Dam_specific)
        # Check if Dam_year is empty
      if (nrow(Dam_year) == 0) {
        print("Dam_year is empty. Breaking the loop!")
        next
      }  
      #_____________________________________________________________________________
      # Downstream
      Downstream_reach_ID_1 <- NHD_Shape %>%
        filter(ID == Dam_specific & GNIS_ID == Dam_NHD_GNIS) %>%
        pull(toID)
      
      # Check if the first downstream reach ID exists
      if (length(Downstream_reach_ID_1) == 0 || Downstream_reach_ID_1 == 0) {
        Downstream_reach_ID_2 <- 0
        Downstream_reach_ID_3 <- 0
      } else {
        # Get the second downstream reach ID
        Downstream_reach_ID_2 <- NHD_Shape %>%
          filter(ID == Downstream_reach_ID_1 & GNIS_ID == Dam_NHD_GNIS) %>%
          pull(toID)
        
        # Check if the second downstream reach ID exists
        if (length(Downstream_reach_ID_2) == 0) {
          Downstream_reach_ID_3 <- 0
        } else {
          # Get the third downstream reach ID
          Downstream_reach_ID_3 <- NHD_Shape %>%
            filter(ID == Downstream_reach_ID_2 & GNIS_ID == Dam_NHD_GNIS) %>%
            pull(toID)
        }
      }
      
      # Create a list of the downstream reach IDs
      downstream_ids <- list(Downstream_reach_ID_1, Downstream_reach_ID_2, Downstream_reach_ID_3)
      
        if (all(unlist(downstream_ids) == 0)) {
          print("All values in downstream_ids are 0. Breaking the loop!")
          next
      }
      # Initialize a list to store the filtered dataframes
      downstream_years <- list()
      
      # Loop over each Downstream reach ID
      for (i in 1:3) {
        # Get the current downstream reach ID
        reach_id <- downstream_ids[[i]]
        
        # If the reach ID is NA, empty, or invalid, create an empty dataframe with the renamed column
        if (is_empty(reach_id) || is.na(reach_id) || reach_id == "") {
          downstream_years[[i]] <- Remote_averaged_filtered[0, ] %>%
            rename(!!paste0("TP_", i, "_Down") := TP)  # Dynamically name the TP column
        } else {
          # Otherwise, filter and rename
          downstream_years[[i]] <- Remote_averaged_filtered %>%
            filter(ID == reach_id) %>%
            rename(!!paste0("TP_", i, "_Down") := TP)  # Dynamically name the TP column
        }
      }
      
      # Assign each downstream year dataframe to the appropriate variable
      Downstream_year_1 <- downstream_years[[1]]
      Downstream_year_2 <- downstream_years[[2]]
      Downstream_year_3 <- downstream_years[[3]]
      
      #_____________________________________________________________________________
      # Join the upstream data by 'Year'
      Upstream_data <- Upstream_year_1 %>%
        left_join(Upstream_year_2, by = "Year") %>%
        left_join(Upstream_year_3, by = "Year") %>%
        select(Year, TP_1_Up, TP_2_Up, TP_3_Up)
      
      Downstream_data <- Downstream_year_1 %>%
        left_join(Downstream_year_2, by = "Year") %>%
        left_join(Downstream_year_3, by = "Year") %>%
        select(Year, TP_1_Down, TP_2_Down, TP_3_Down)
      
      Final_data <- Dam_year %>%
        left_join(Upstream_data, by = "Year") %>%
        left_join(Downstream_data, by = "Year")
      
      Dam_result <- Final_data %>%
        select(Year, TP_1_Up, TP_2_Up, TP_3_Up, TP, TP_1_Down, TP_2_Down, TP_3_Down)
      
      ordered_TP_Type <- c('TP_3_Up', 'TP_2_Up', 'TP_1_Up', 'TP', 'TP_1_Down', 'TP_2_Down', 'TP_3_Down')
      
      medians <- apply(Dam_result, 2, median, na.rm = TRUE)
      
      # Convert the result into a single-row DataFrame
      medians_df <- as.data.frame(t(medians))
      
      medians_df <- medians_df%>%mutate("Dam_height" = Dam_height, "Count" = nrow(Dam_result))
      
      # Assign names to the last two columns
      colnames(medians_df)[(ncol(medians_df)-1):ncol(medians_df)] <- c("Dam_height", "Count") 
      
      # Set column names to match the original dataset
      colnames(medians_df) <- colnames(Dam_result)
      
      result <- rbind(result, medians_df)

      
}
```
#Dam TP change
```{r}
# Assign names to the last two columns
colnames(result)[(ncol(result)-1):ncol(result)] <- c("Dam_height", "Count") 

ordered_TP_Type <- c('TP_3_Up', 'TP_2_Up', 'TP_1_Up', 'TP', 'TP_1_Down', 'TP_2_Down', 'TP_3_Down')

# Filter the data based on dam height and number of yearly observations
result_filtered <- result%>%filter(Dam_height > 30.5, Count>= 15)

result_filtered$Upstream <- rowMeans(result_filtered[, c("TP_1_Up", "TP_2_Up", "TP_3_Up")], na.rm = TRUE)
result_filtered$Downstream <- result_filtered$TP

# Calculate the average for each column reach
result_filtered_average <- apply(result_filtered, 2, mean, na.rm = TRUE)

result_filtered_average 

# Extract the average values for upstream columns and calculate their mean
Upstream_TP <- mean(
  c(
    result_filtered_average["TP_1_Up"], 
    result_filtered_average["TP_2_Up"], 
    result_filtered_average["TP_3_Up"]
  ), 
  na.rm = TRUE
)

# Extract the average value for the downstream column
Downstream_TP <- result_filtered_average["TP"]

# Calculate percent change
Percent_change <- ((Upstream_TP - Downstream_TP) / Downstream_TP) * 100

# Print the result
print(paste("Overall Percent change was:", Percent_change))
```
# Wilcoxin Test
```{r}
# Perform the Wilcoxon test
wilcox_test <- wilcox.test(result_filtered$Upstream, result_filtered$Downstream, paired = TRUE)

# Check if the result is significant and print the appropriate message
if (wilcox_test$p.value < 0.05) {
  print("Downstream and Upstream reaches are statistically different.")
} else {
  print("No statistically significant difference between downstream and upstream reaches.")
}

```

# Considering only instances where the downstream TP was less than the upstream TP
```{r}
# Filter rows where TP_1_Up > TP
result_filtered_positive <- result_filtered %>%
  filter(TP_1_Up > TP)

# Calculate the average for each column and store it as a named vector
result_filtered_positive_average <- colMeans(result_filtered_positive, na.rm = TRUE)

# Extract the average values for upstream columns and calculate their mean
Upstream_TP <- median(
  c(
    result_filtered_positive_average["TP_1_Up"], 
    result_filtered_positive_average["TP_2_Up"], 
    result_filtered_positive_average["TP_3_Up"]
  ), 
  na.rm = TRUE
)

# Extract the average value for the downstream column
Downstream_TP <- result_filtered_positive_average["TP"]

# Calculate percent change
Percent_change <- ((Upstream_TP - Downstream_TP) / Downstream_TP) * 100

# Print the result
print(paste("Percent change in the positive was:", Percent_change))
```

# Considering only instances where the downstream TP was more than the upstream TP
```{r}
# Filter rows where TP_1_Up > TP
result_filtered_negative <- result_filtered %>%
  filter(TP_1_Up < TP)

# Calculate the average for each column and store it as a named vector
result_filtered_negative_average <- colMeans(result_filtered_negative, na.rm = TRUE)

# Extract the average values for upstream columns and calculate their mean
Upstream_TP <- median(
  c(
    result_filtered_negative_average["TP_1_Up"], 
    result_filtered_negative_average["TP_2_Up"], 
    result_filtered_negative_average["TP_3_Up"]
  ), 
  na.rm = TRUE
)

# Extract the average value for the downstream column
Downstream_TP <-result_filtered_negative_average["TP"]

# Calculate percent change
Percent_change <- ((Upstream_TP - Downstream_TP) / Downstream_TP) * 100

# Print the result
print(paste("Percent change in the positive was:", Percent_change))
```
#Make the plot
```{r}
Percent_change <- result_filtered

# Calculate the mean of upstream columns for each row
Percent_change$Upstream <- rowMeans(
  cbind(Percent_change$TP_1_Up, Percent_change$TP_2_Up, Percent_change$TP_3_Up),
  na.rm = TRUE
)

# Downstream column is directly assigned
Percent_change$Downstream <- Percent_change$TP

# Calculate percent change for each row
Percent_change$Change <- ((Percent_change$Upstream - Percent_change$Downstream) / 
                          Percent_change$Downstream) * 100


Dam_plot <- ggplot(Percent_change, aes(x = Change, fill = Change > 0)) +
  geom_histogram(alpha = 0.8, , position = "identity") +  # Histogram with transparency and 30 bins
  scale_fill_manual(values = c("lightcoral", "lightblue"),  # Color for negative and positive
                    name = "Change Direction", 
                    labels = c("Negative", "Positive")) +  # Label the legend
  labs(title = "Dam", x = "Change (%)", y = "Count") +  # Adjust labels
  theme_minimal() +  # Use a clean minimal theme
  scale_y_continuous(expand = c(0, 0)) +  # Adjust y-axis scaling
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),  # Center and size the title
    axis.title = element_text(size = 14),  # Increase axis title size
    axis.text = element_text(size = 14),  # Increase axis text size
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(size = 1),  # Increase axis line thickness
    legend.title = element_text(size = 14),  # Increase legend title size
    legend.text = element_text(size = 14),
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add a box around the plot
  )+ylim(0,25)+xlim(-100, 150) # Adjust x-axis limits


Dam_plot
ggsave("C:\\Research\\TP Analysis\\Figures\\Supporting Figures\\Dam.png", plot = Dam_plot , width = 10, height = 8, units = "in", dpi = 1000)
```




